Optogenetic Maxwell Demon to Exploit Intrinsic Noise and Control Cell Differentiation Despite Time Delays and Extrinsic Variability

### 2023-04-26 paper.tex
update

\begin{abstract}
Synthetic biology seeks to create modular components that generate complex and controllable biological behaviors, often at microscopic scales. In single-cell applications, where important regulatory RNA or proteins are present in small numbers, intrinsic fluctuations cause biochemical `noise' that frequently results in undesired effects, and this noise is usually treated as a nuisance that must be removed. However, recent theoretical analyses have shown that noise can sometimes be used to one's advantage when seeking to manipulate nonlinear regulatory systems that would be difficult or impossible to control in a fully deterministic setting. Specifically, it was shown that a noise-exploiting optogenetic controller reminiscent of Maxwell's Demon could systematically force multiple identical systems to reach different independently-assigned stable points, even for a few special cases where the model is inexact or if observations were incomplete. This work expands upon those analyses to explore in greater detail how control performance is affected by errors or approximations in the model or by temporal delays in the observations. We find that noise-exploiting controllers can remain highly effective despite coarse approximations to the model's scale or incorrect estimations of key model parameters, and these controllers can even retain performance for significant time delays.  Together, these findings suggest that noise-exploiting control should be possible in real experiments where models are always approximate, where parameters are always uncertain, and where observations are corrupted by errors.

@@

\section{Introduction}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{CartoonAndControler.pdf}
\caption{{\bf Optogenetic control of multiple cells using a single light course}
({\bf A}) Schematic of the gene expression system. UV light activates T7 polymerase based transcription and protein expression from the first promoter-gene construct, and the protein actuates its own production via a second promoter-gene construct.
({\bf B}) Schematic of the optogenetic control system. An observer is partially (PAC) or fully (FAC) aware of cell expression. The controller modulates a single spatially homogeneous light signal in attempt to drive each genetically identical cell to its own specified state.  
({\bf C}) The FAC (top) and PAC (bottom) control laws for two cells ($\mathbf{u}^{XAC}(x_1,x_2)$ shown on left) break the symmetry between Cell 1 and Cell 2 and result in distinct marginal distributions for the cells' responses (right).
Figure adapted from \cite{May2021}.}
\label{cartoons}
\end{center}
\vspace{-0.3in}
\end{figure}
% XXX - Please modify figure as follows: (1) switch panels A and B to match caption. (2) Change title of panel B (previously panel A) to "External Observer, controller and actuator".  (3) Change title of panel A (previously panel B) to "Internal gene regulation pathway". (4) Change title of panel C to "Feedback control law and resulting performance"
% MMM - Made the edit

Synthetic biology seeks to create modular\cite{Ng2019} and orthogonal\cite{Liu2018} components to sense and actuate\cite{Sheets2020} complex logical systems\cite{Groseclose2020}, and which are capable of performing a variety of advanced biological behaviors\cite{Shin2020}. 
New optogenetic tools have increased our ability to actuate embedded systems within cells reliably and with strong control performance\cite{Sheets2020,Baumschlager2017,Chen2020,Lillacci2018}. Great advances in these two fields have enabled computer programmable control of cellular protein production through the use of external optogenetic inputs and smart microscope techniques\cite{Fox2021,Baumschlager2021,Lugagne2017}. These types of digital-synthetic actuators allow for fine-tuned, computer-modulated control of cellular systems that were previously impossible\cite{Rullan2018, Baumschlager2017} and with fast response times in comparison to chemical diffusion.

@@

\section{Methods}
In previous work~\cite{May2021}, a six-species gene regulation model was postulated for the light-activated association of two T7 split domains (species 1 and 2) to form active T7 polymerases (species 3) which could then in turn associate with inactive T7 promoters (species 4) to create an active gene construct (species 5) that would produce its eventual protein product (species 6). A second, far simpler single-species model was also postulated by assuming quasi-steady equilibrium for the first five species. Both models were independently parametrized by fitting to the same experimental data from Baumschlager, et al.~\cite{XXX}, and then extended to include a secondary self-activated promoter-gene construct for which the expression rate was defined by a Hill function (see Eqn.\ \ref{prodRate} below). In \cite{May2021}, we showed that a control law identified using the simple model would work to achieve differential control for the more complex model, thus motivating our current study's efforts to explore further how control performance would change upon additional simplifying assumptions or errors in model specification.  In the following subsections, we discuss our methods to formulate this model, show how it is analyzed numerically, and discuss the specification of the control objective and the determination of an optimal controller.

@@

To assess the potential robustness of noise-enhanced control of the proposed synthetic auto-repression toggle switch, we adopt the model and  parameters from \cite{May2021} as well as the fully aware and partially aware controllers, $u^{\rm FAC}(x_1,x_2)$ and  $u^{\rm FAC}(x_1)$, respectively as described in the methods section.  In the following subsections, we explore how well these controllers would perform in the more realistic setting where the approximate model used to define the controller no longer matches exactly to the dynamics of the process that the controller is being used to manipulate.

@@

To shed light on this tradeoff, control performance scores for the $\mathbf{u}^{\rm FAC}$ and $\mathbf{u}^{\rm PAC}$ controllers were calculated at different levels of granularity ($\alpha$) between 0.1 and 2.0. 
% XXX - Why only go up to 2?  Did it become intractable?
% MMM - Non_Sparse Calculations are intractible after 2, but sparse versions of code will work fine
Figure \ref{Volume} (Ai - Fi) shows the joint (left plots) and marginal (right plots) distributions of the two cells relative to the specified target position (circles in joint distributions and vertical dashed lines in marginal distributions) for the FAC controller and for $\mathbf{\alpha}$=.2 (A,B), 1 (C,D) and 2 (E,F). As the granularity changes from $\alpha$=0.2 , to $\alpha$=2, the distributions become focused at a point near the target, and the performance score improves from $J=$333 to 85 to 27, respectively. The PAC controller performance also improves considerably with alpha as shown in Fig.\ \ref{Volume} (G-L).
% XXX - Please change the panel labels to A-F, G-L, and M.
% MMM - made the change
\subsection{Effects of changes to system granularity}
\begin{figure*}[t!]
\begin{center}
\includegraphics[width=1\textwidth]{GranularityPerturbation.eps}
\vspace{-0.1in}
\caption{Effects of Scale Changes. (A-F) Joint (A,C,D,) and marginal (B,D,F) probability distributions for the FAC control method at $\alpha=0.2$ (A,B), $\alpha=1$ (C,D) and $\alpha=2.0$ (E,F).  Target point shown as blue circle. (G-L) Same as (A-F), but for PAC controller. (M)  FAC (cyan) and PAC (magenta) controller performance as function of system volume. Horizontal dashed lines correspond to SSA analyses at $\alpha=100$. In all analyses, the system model and PAC and FAC controllers are unchanged from \cite{May2021} other than their adaptation to the new system volume size (see Methods).}
\label{Volume}
\end{center}
\vspace{-0.2in}
\end{figure*}
% XXX - This figure needs larger fonts for all the labels and axes. 
% MMM - made the change
% Update panel letters as written in caption
% Add tic marks to left side of the A and C heat maps

@@

\begin{figure*}
\begin{center}
\includegraphics[width=1\textwidth]{GranularityScore.eps}
\caption{}
\label{}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1\textwidth]{CalculationSpeed.eps}
\caption{}
\label{Speed}
\end{center}
\end{figure*}

We note however, that the improved steady state performance for systems with larger scales, comes at a cost that they also take longer to reach a steady state distribution. This tradeoff is illustrated in Fig.\ \ref{Score} which shows the performance score $J(t)$ as the system adapts to a change from the original target point $\mathbf{T} = [30,10]$ to its mirror point $\mathbf{T}_m = [10,30]$.  
% XXX - Please add a new figure to show this point.  Start with an initial condition that is the steady state with target T and then flip it to T' = T(::-1).  Then solve for P(t) and calculated J(t) from this.    Finally, make a plot of J(t) versus t for the three cases \alpha = 0.2, 1, 2.
One hundred SSA simulations of the flip-mirroring process were simulated while measuring the average score all simulations over time. Fig.\ \ref{Score} (A and B) shows the average score and protien counts of the simulation when $\alpha=1$, Fig. \ \ref{Score} (C,D) when $\alpha = 10$, and Fig. \ \ref{Score} (E,F) when $\alpha = 100$.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=1\textwidth]{ParameterPerturbation.pdf}
\caption{Effects of parametric variations on control performance.  ({\bf A}) Sensitivity of control performance score, $J$, to modifications in parameters applied to Cell 1 only, one at a time for parameters $\kappa$, $\beta$, $\eta$, $k_0$, and $\gamma$ (in order left to right).  ({\bf B}) Same as {\bf A}, but where modifications are made to the corresponding parameter in Cell 2. ({\bf C}) Same as ({\bf A}), but where the modification is made simultaneously to both cells. In all panels, trends lines are shown for the FAC (solid) and PAC (dashed) controllers. Controllers are unchanged from \cite{May2021}.}
\label{Parameter}
\end{center}
\end{figure*}
% XXX - This figure needs much larger fonts for all the labels and axes.
% MMM - Did the change
% There are errors in the parameter names.  What is 'alpha'?
% remove the A,B,C,D,E.
% replace I, II, III with A, B, C
% what is parameter $\mu$? Is that an error?  Please use $\mu$ for reaction index as in the methods section.

\subsection{Effects of Parameter Errors or Extrinsic Uncertainties}

Exact parameters are never known for any real biological process, and even genetically identical single-cell systems are expected to exhibit cell-to-cell heterogeneity in parameters due to extrinsic variations outside of the specific pathways of interest. These parameter uncertainties or variations between cells could influence system performance, such that a control law that is optimized for one fixed, but incorrect, set of parameters may or may not succeed if used to break symmetry and track a desired target state when applied to a system with different parameters. To explore the sensitivity of control performance to unknown errors in the model parameters, sensitivity analysis was performed for each model parameter and for the situation where the corresponding parameter was modified in either one cell at a time (i.e., to simulate extrinsic noise) or in both cells simultaneously (i.e., to simulate parameter errors).  For each perturbation analysis, both the $\mathbf{u}^{\rm FAC}$ and $\mathbf{u}^{\rm PAC}$ controller were analyzed over a range of parameters from one-half to two-fold of the original parameters.
% XXX - How hard would it be to extend this to 10 fold in either direction?
% MMM-Changes to parameters can dramatically change the location of the probability distribution which is contained in a bounded (50,50) area. A 10 fold increase to a parameter can push all the probability to the top corner of the space
 Figure \ref{Parameter}A shows the results for these parameter sweeps when a single parameter in Cell 1 is change while the parameters of Cell 2 is held fixed, and Fig.\ \ref{Parameter}B shows the opposite when changes Cell 1 has fixed parameters and Cell 2 is modified.  Finally, \ref{Parameter}B  shows the performance change when the corresponding parameter is changed in both Cell 1 and Cell 2 at the same time.

\subsection{Effects of Time Delays on Controller Performance}

\begin{figure*}
\begin{center}
\includegraphics[width=1\textwidth]{TimeDelayPerturbation.eps}
\vspace{-0.1in}
\caption{Effects of time delays on control performance. (A-F) Joint (A,C,E) and marginal (B,D,F) probability distributions for the FAC control method at three different time delays of $\tau/\gamma$=0.01 (A,B), 0.1 (C,D) and 1.0 (E,F) relative to the target point (blue circles). (G-L) Same as (A-F), but for PAC controller. (M) FAC and PAC performance as function of time delay show that PAC outperforms the FAC when time delay is greater than $0.07 /\gamma$. In all analyses, the system model and PAC and FAC controllers are unchanged from \cite{May2021}; specifically, the protein degradation rate is $\gamma = 0.0203$ as identified from \cite{Baumchager}.}
\label{Time}
\end{center}
\vspace{-0.2in}
\end{figure*}
% XXX - This figure needs larger fonts for all the labels and axes. 
% MMM- Did the change
% Update panel letters as written in caption
% Add tic marks to left side of the A and C heat maps
% Fix the ref to baumschlager paper.

@@

>>>>

\begin{abstract}

Unnderstanding the effects of noise in control theory is important to ()
Much work has been done to make control performance robust to noise, but the development of controllers which use noise to their advantage is not fully understood.
Motivated by optogentics, we proposed a difficult control problem which requires the exploitation of noise in a stochastic system to break symmetry between two signals.
in previous work, we found multiple such controllers which could exploit noise to break symmetry between two cells under a variety of system information.
This work extends that analysis to include stochastic systems that trace a moving target, are affected by time delay, intrinsic noise, or measurement uncertainty.

Synthetic biology seeks to create modular components that generate complex and controllable biological behaviors, often at microscopic scales where important regulatory RNA or proteins are present in small numbers.  Intrinsic fluctuations at these scales cause biochemical `noise' that frequently results in undesired neussance effects. However, recent theoretical analyses have shown that noise can sometimes be used to one's advantage when seeking to manipulate nonlinear regulatory systems that would be difficult or impossible to control in a fully deterministic setting. Specifically, it was shown that a noise-exploiting optogenetic controller reminiscent of Maxwell's Demon could systematically force multiple identical systems to reach different independently-assigned stable points, even for a few special cases where the model is inexact or if observations were incomplete. This work expands upon those analyses to explore in greater detail how control performance is affected by errors or approximations in the model or by temporal delays in the observations. We find that noise-exploiting controllers can remain highly effective despite coarse approximations to the model's scale or incorrect estimations of key model parameters, and these controllers can even retain performance for significant time delays.  Together, these findings suggest that noise-exploiting control should be possible in real experiments where models are always approximate, where parameters are always uncertain, and where observations are corrupted by errors.

@@

\section{Introduction}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{CartoonAndControler.pdf}
\caption{{\bf Optogenetic control of multiple cells using a single light input.}
({\bf A}) Schematic of the light activated genetic system with autoregulation.
({\bf B}) Diagram of the stochastic SIMO control problem using two optogenetic cells sharing an input.
({\bf C}) Adapted from (PAPER). Noise exploiting controllers were optimized to a fully aware control input (I), and partially aware control input(III). Marginal distributions are solved using the finite state projection and show a break in symmetry using both controllers (II and IV) }
\label{cartoons}
\end{center}
\vspace{-0.3in}
\end{figure}
% XXX - Please modify figure as follows: (1) switch panels A and B to match caption. (2) Change title of panel B (previously panel A) to "External Observer, controller and actuator".  (3) Change title of panel A (previously panel B) to "Internal gene regulation pathway". (4) Change title of panel C to "Feedback control law and resulting performance"
% MMM - Made the edit

Synthetic biology seeks to create modular\cite{Ng2019} and orthogonal\cite{Liu2018} components to sense and actuate\cite{Sheets2020} complex logical systems\cite{Groseclose2020}, and which are capable of performing a variety of advanced biological behaviors\cite{Shin2020}. 
New optogenetic tools have increased our ability to actuate embedded systems within cells reliably and with strong control performance\cite{Sheets2020,Baumschlager2017,Chen2020,Lillacci2018}. Advances in these two fields have enabled computer programmable control of cellular protein production through the use of external optogenetic inputs and smart microscope techniques\cite{Fox2021,Baumschlager2021,Lugagne2017}. These types of digital-synthetic actuators allow for fine-tuned, computer-modulated control of cellular systems that were previously impossible\cite{Rullan2018, Baumschlager2017} and with fast response times in comparison to chemical diffusion.

@@

\section{Methods}
In previous work~\cite{May2021}, a six-species gene regulation model was developed to describe light-activated association of two T7 split domains (species 1 and 2) to form active T7 polymerases (species 3) which could then in turn associate with inactive T7 promoters (species 4) to create an active gene construct (species 5) that would produce its eventual protein product (species 6). A second, far simpler single-species model was also developed by assuming quasi-steady equilibrium for the first five species. Both models were independently parametrized by fitting to the same experimental data from Baumschlager, et al.~\cite{XXX}, and then extended to include a secondary self-activated promoter-gene construct for which the expression rate was defined by a Hill function (see Eqn.\ \ref{prodRate} below). In \cite{May2021}, we showed that a control law identified using the simple model would work to achieve differential control for the more complex model, thus motivating our current study's efforts to explore further how control performance would change upon additional simplifying assumptions or errors in model specification.  In the following subsections, we discuss our methods to formulate this model, show how it is analyzed numerically, and discuss the specification of the control objective and the determination of an optimal controller.

@@

\subsection{Effects of Parameter Errors or Extrinsic Uncertainties}
\begin{figure*}
\begin{center}
\includegraphics[width=1\textwidth]{ParameterPerturbation.pdf}
\caption{Parameter sweeps using the FAC and PAC show a broad range of control performance in cell 1 ({\bf A}),  in cell 2 ({\bf B}), and in both cells ({\bf C}). }
\label{Parameter}
\end{center}
\end{figure*}

Exact parameters values are typically unknown for real stochastic process, and even genetically identical single-cell systems can exhibit cell-to-cell heterogeneity in parameters due to extrinsic variations outside of the specific pathways of interest. Estimates of parameters can be taken, but not exact measurement. These parameter uncertainties could influence system performance. To explore the sensitivity of control performance to unknown errors in the model parameters, sensitivity analysis was performed for each model parameter and for the situation where the corresponding parameter was modified in either one cell at a time (i.e., to simulate extrinsic noise) or in both cells simultaneously (i.e., to simulate parameter errors).  For each perturbation analysis, both the $\mathbf{u}^{\rm FAC}$ and $\mathbf{u}^{\rm PAC}$ controller were analyzed over a range of parameters from one-half to two-fold of the original parameters.
% XXX - How hard would it be to extend this to 10 fold in either direction?
% MMM-Changes to parameters can dramatically change the location of the probability distribution which is contained in a bounded (50,50) area. A 10 fold increase to a parameter can push all the probability to the top corner of the space
 Figure \ref{Parameter}A shows the results for these parameter sweeps when a single parameter in Cell 1 is change while the parameters of Cell 2 is held fixed, and Fig.\ \ref{Parameter}B shows the opposite when changes Cell 1 has fixed parameters and Cell 2 is modified.  Finally, \ref{Parameter}B  shows the performance change when the corresponding parameter is changed in both Cell 1 and Cell 2 at the same time.

To assess the potential robustness of noise-enhanced control of the proposed synthetic auto-repression toggle switch, we adopted the model  from \cite{May2021} as well as the fully aware and partially aware controllers, $u^{\rm FAC}(x_1,x_2)$ and  $u^{\rm FAC}(x_1)$, respectively as described in the methods section.  In the following subsections, we explore how well these controllers would perform in the more realistic setting where the approximate model used to define the controller no longer matches exactly to the dynamics of the process that the controller is being used to manipulate.

@@

To shed light on this tradeoff, control performance scores for the $\mathbf{u}^{\rm FAC}$ and $\mathbf{u}^{\rm PAC}$ controllers were calculated at different levels of granularity ($\alpha$) between 0.1 and 2.0. 
% XXX - Why only go up to 2?  Did it become intractable?
% MMM - Non_Sparse Calculations are intractible after 2, but sparse versions of code will work fine
Figure \ref{Volume} (Ai - Fi) shows the joint (left plots) and marginal (right plots) distributions of the two cells relative to the specified target position (circles in joint distributions and vertical dashed lines in marginal distributions) for the FAC controller and for $\mathbf{\alpha}$=.2 (A,B), 1 (C,D) and 2 (E,F). As the granularity changes from $\alpha$=0.2 , to $\alpha$=2, the distributions become focused at a point near the target, and the performance score improves from $J=$333 to 85 to 27, respectively. The PAC controller performance also improves considerably with alpha as shown in Fig.\ \ref{Volume} (G-L).
% XXX - Please change the panel labels to A-F, G-L, and M.
% MMM - made the change
\subsection{Effects of changes to system granularity}
\begin{figure*}[t!]
\begin{center}
\includegraphics[width=1\textwidth]{GranularityPerturbation.pdf}
\vspace{-0.1in}
\caption{Systems with high granularity were shown to have increased control performance but slower response times.}
\label{Volume}
\end{center}
\vspace{-0.2in}
\end{figure*}
% XXX - This figure needs larger fonts for all the labels and axes. 
% MMM - made the change
% Update panel letters as written in caption
% Add tic marks to left side of the A and C heat maps

@@

We note however, that the improved steady state performance for systems with larger scales, comes at a cost that they also take longer to reach a steady state distribution. This tradeoff is illustrated in Fig.\ \ref{Score} which shows the performance score $J(t)$ as the system adapts to a change from the original target point $\mathbf{T} = [30,10]$ to its mirror point $\mathbf{T}_m = [10,30]$.  
% XXX - Please add a new figure to show this point.  Start with an initial condition that is the steady state with target T and then flip it to T' = T(::-1).  Then solve for P(t) and calculated J(t) from this.    Finally, make a plot of J(t) versus t for the three cases \alpha = 0.2, 1, 2.
One hundred SSA simulations of the flip-mirroring process were simulated while measuring the average score all simulations over time. Fig.\ \ref{Score} (A and B) shows the average score and protein counts of the simulation when $\alpha=1$, Fig. \ \ref{Score} (C,D) when $\alpha = 10$, and Fig. \ \ref{Score} (E,F) when $\alpha = 100$.

@@

\subsection{Effects of Time Delays on Controller Performance}
\begin{figure*}
\begin{center}
\includegraphics[width=1\textwidth]{TimeDelayPerturbation.pdf}
\vspace{-0.1in}
\caption{Stochastic simulations of the time delayed system show decreasing control performance using either controller. }
\label{Time}
\end{center}
\vspace{-0.2in}
\end{figure*}
% XXX - This figure needs larger fonts for all the labels and axes. 
% MMM- Did the change
% Update panel letters as written in caption
% Add tic marks to left side of the A and C heat maps
% Fix the ref to baumschlager paper.

@@

% XXX - This figure needs much larger fonts for all the labels and axes.
% MMM - Did the change
% There are errors in the parameter names.  What is 'alpha'?
% remove the A,B,C,D,E.
% replace I, II, III with A, B, C
% what is parameter $\mu$? Is that an error?  Please use $\mu$ for reaction index as in the methods section.

\begin{figure*}
\begin{center}
\includegraphics[width=1\textwidth]{DelayAndGranularity.pdf}
\vspace{-0.1in}
\caption{Stochastic simulations of the FAC driven system (A) and PAC driven system (B) show different failure modes at at different $(\tau,\alpha)$ pairs (C).}
\label{DG}
\end{center}
\vspace{-0.2in}
\end{figure*}

Joint analysis of time delay and granularity together was performed to better understand if granularity can undo the effect of time delay. System control performance was analyzed over a two dimensional domain of points $(\tau,\alpha)$ using sixteen stochastic simulations for XXX minutes. These simulations were driven using the FAC controller Fig \ref{DG} (A), and PAC controller Fig \ref{DG} (B). The landscape of the FAC and PAC control performance Fig \ref {CtrlP}(A and B)over the domain shows a valley at low $\tau$, and a plateau at large $\tau$. This suggests that increasing system scale at large time delays is not beneficial to control performance. To better understand the appearance of the landscape, trajectories in different regions were compared. Trajectories at the edge of the plateau region (blue star) showed bursty behavior Fig \ref {CtrlP}(C)(blue lines), while trajectories at the edge of the valley Fig \ref {CtrlP}(C) (red lines) showed oscillatory behavior. Despite the small different in time delay between the red and blue trajectory, a stark change to the overall behavior was observed.

\begin{figure*}
\begin{center}
\includegraphics[width=1\textwidth]{ControlPoints.pdf}
\vspace{-0.1in}
\caption{Controllers optimized to a 2d domain of target points (A) show good control performance in the central region despite a broad range of targets (B and C). }
\label{CtrlP}
\end{center}
\vspace{-0.2in}
\end{figure*}

Many controllers were optimized to different target points to determine if certain reference points were harder to control than others. Controllers were optimized over a two dimensional domain of target points $(T_1,T_2)$between between five and fourty-five. Figure \ref{CtrlP} (A) shows the control performance of the system after optimizing each controller to each point in the domain. Figure \ref{CtrlP} (B and C) show the optimized control input and the steady state of two points in the system. The unstable point in the rate equation matches the unstable point near $(20,20)$ in this domain.

\begin{figure*}
\begin{center}
\includegraphics[width=1\textwidth]{ControlReference.pdf}
\vspace{-0.1in}
\caption{Finite state projections of the slowly driven system demonstrate good control performance under a variety of phase and frequency scenario. }
\label{CR}
\end{center}
\vspace{-0.2in}
\end{figure*}

Dynamic reference targeting was performed by cycling though individual controllers which were optimized to a single target point at a fixed frequency. Thirty-two controllers were optimized along a pathway in order to target an in-sync reference point \ref{CR}(Synchronous), a phase lagged reference point \ref{CR}(Phase Lagged), or frequency separated reference point\ref{CR}(Phase Lagged and Frequency). 
The in phase signal is given by $r(t)=[10 sin (2\pi f t)+20, 10 sin (2\pi f t)+20]$, the phase-lagged signal is given by $r(t)=[10 sin (2\pi f t)+20, 10 cos (2\pi f t)+20]$, and the frequency driven signal is given by $r(t)=[10 sin (2\pi f t)+20, 10 sin (3\pi f t)+20]$ FSP simulations of cyclically driven systems were slowly driven at a frequency of 1/5000 seconds and their medians are shown in the Fig \ref{CR}(A4, B4,C4)(red cloud, blue cloud). Root mean squared error of each simulation are shown in Fig \ref{CR}(A5, B5,C5). Simulations show that in-phase control performed the best with an RSME of 7.0 despite moving over the unstable point in the system. When the system is driven with a phase-lag the RSME of the score increases to 8.3 and when driven at a different frequency the RMSE goes up to $9.1$.

\begin{figure*}
\begin{center}
\includegraphics[width=1\textwidth]{ControlReferenceGranularity.pdf}
\vspace{-0.1in}
\caption{Stochastic simulations driven using a phased lagged controller show improved score at high granularity and low frequency.}
\label{CRG}
\end{center}
\vspace{-0.2in}
\end{figure*}
Although the control of a dynamic reference point using a Piecewise controller is possible, the effect of cycling frequency is also important.
16 SSA simulations of the phase-separated driving controller were performed on a system over a two dimensional domain of points $(\alpha,f)$ for a time of XXX minutes. Comparisons of control performance at $\alpha=0.5$ and $\alpha=2.0$ (Fig \ref}{CRG} (C))show that control performance improves with system scale. 16 SSA simulations of the high granularity system driven by phase-lag were performed over a range of frequency values.  These analysis show that at low frequency the systems have a noise floor which decreases with increasing alpha, but was the system were driven rapidly, all systems became equally uncontrollable. At high frequency, the scores never got worse than $J=650$. At high frequency the control performance plateaus, while at lower frequency the system improves with increasing $\alpha$ Fig \ref{CRG}(E).
